---
title: "Data Tidying"
author: "Monika Sikand"
date: "1/16/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning = F, message = F}
library(dplyr)
library(tidyr)
#library(stats)
```
 to call a function from a specific package 'package_name::function_name(...)'
 
# Data Cleaning

Read in data file
```{r}
catch_df <- read.csv(url( "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",method = "libcurl"), stringsAsFactors = FALSE)
head(catch_df)
```

control +SHift + m gives %>% pipe operator
or

command + shift + m gives %>% pipe operator

catch_cleaned <- catch_df %>% 
  select(-All, -notesRegCode)
  
  * Remove marginal some and notes column
  * Move from wide to long format

```{r}
catch_long <- catch_df %>% 
  select(-All, -notesRegCode)%>%  # negative values remove columns
  gather(key = "species", value = "catch",-Year, -Region)

head(catch_long)
```

%>% 

```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch)

head(catch_cleaned)
```
* erroneus value due to OCR issue - chnage "I" to one
* create catch column in correct units


```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands))%>%  
  mutate(catch_thousands = as.integer(catch_thousands)) %>%         
  mutate(catch = catch_thousands * 1000)
 
 tail(catch_cleaned)
```

```{r, eval = F, echo = F}
# I used this code to find the bad value. 
test_catch <- as.integer(catch_cleaned$catch_thousands)

i <- which(is.na(test_catch) == T)

catch_cleaned[i, ]
```
# Split- Apply-Combine

Calculate total catch by region

```{r}
catch_total <- catch_cleaned%>%
  group_by(Region, Year) %>% 
  summarize(catch_region = sum(catch),
                   n_obs = n(), 
                  n_mean = mean(catch),
                  n_median = median(catch))

catch_total
```
Matt's challenge

```{r}
catch_matt <- catch_cleaned %>%
   group_by(species, Year) %>%
  summarize(catch_mean = mean(catch),# calculate mean
           catch_sd = sd(catch),
           catch_n  = n()) #total observations
head(catch_matt)
```


Filter for Chinook Salmon
 # | is or operator
 # & is and operator

```{r}
catch_chinook <- catch_cleaned %>%
  filter(species == "Chinook" & Region == "SSE" & Year > 1990)%>%
  arrange(Year) # sorting in descending order
head(catch_chinook)
```

* select
* filter
* mutate
* rename
* gather
* spread
* arrange

# Joins

Read in region defs

```{r}

region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)

head(region_defs)
```


```{r}
region_clean <- region_defs%>%
  select(code, mgmtArea)
head(region_clean)

```


```{r}

catch_joined <- left_join (catch_cleaned, region_clean, 
                          by = c("Region" = "code"))
  
head(catch_joined)
```


