---
title: "index"
author: "Monika Sikand"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```


# Data Tyding

Escapement data are source from: Alaska Department of fish and Game, Division of Commercial Fisheries. 

```{r}
esc <- read.csv(url ("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e" , method = "libcurl"), stringsAsFactors = F)
head(esc)

```

calculate annual, regional total escapement
* `separate` the date into year month date
* `group-by` `summarize` to calculate the sum over the unique region-year-species combinations
* `filter`for salmon species

```{r, echo = F}
annual_esc <- esc %>%
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.numeric(year)) %>%
  group_by(SASAP.Region,Species,year)%>%
  summarise(escapement = sum(DailyCount))%>%
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))
  
head(annual_esc)
```


```{r}
datatable(annual_esc) 
```


# Static Plots

# geom_col(fill = "blue") Use help and cheatsheets. Chosoe data visualization with 
```{r}
ggplot(annual_esc, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col() 
  
```

# time series of Kodiak escapement by species
```{r}

ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
       aes(x = year, y = escapement, color = Species)) +
geom_line() +
  geom_point() + 
  ylab("Escapement (number of fish)") +
ggtitle("Kodiak Salmon Escapement") + 
theme_bw() +
theme(plot.title = element_text(hjust = 0.5))  

```
plot time series  of escapement by species for all regions

```{r, fig.height=10, fig.width=6}
plot_name <- ggplot(annual_esc,aes(x = year, y = escapement, color = Species))+
  geom_line()+
  geom_point()+
  facet_wrap(~SASAP.Region, scales = "free", ncol = 2) +
  theme_bw()
plot_name

ggsave("region_escapement.png", 
       plot = plot_name, 
       height = 10, 
       width = 8, 
       units = "in")

```


# Interactive 